---
title: "FDSC 403 Graphs"
output: html_notebook
author: Jess Strait
---


```{r}
# Front Matter
rm(list = ls())

library(plotrix)
library(ggplot2)
library(dplyr)
line <- read.csv("LineScaleClean.csv")
```

```{r, fig.width=10,fig.height=8}
# Q1: W/V Concentration v. line scale rating
ggplot(line, aes(x=Conc, y=LS, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), size=1, alpha=.2) + labs(title="Weight/Volume Sucrose Concentration Versus Line Scale Rating", x="Weight/Volume Concentration", y="Line Scale Rating (cm)", color="Sweetener", shape="Sweetener") + scale_x_continuous(breaks=c(.025, .05, .1, .15, .2), labels=c(".025", ".05", ".10", ".15", ".20")) + theme_bw()
```

```{r, fig.width=10,fig.height=8}
# Q2: W/V Concentration v. magnitude estimation
ggplot(line, aes(x=Conc, y=ME, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=MEMean-MESE, ymax=MEMean+MESE), size=1, alpha=.2) + labs(title="Weight/Volume Sucrose Concentration Versus Magnitude Estimation", x="Weight/Volume Concentration", y="Magnitude Estimation", color="Sweetener", shape="Sweetener") + scale_x_continuous(breaks=c(.025, .05, .1, .15, .2), labels=c(".025", ".05", ".10", ".15", ".20")) + scale_y_continuous(breaks=c(1, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120))  + theme_bw()
```

```{r, fig.width=12,fig.height=8}
#Q3: Molarity v. line scale rating
ggplot(line, aes(x=Mol, y=LS, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), size=1, width=.05, alpha=.2) + labs(title="Molarity Versus Line Scale Rating", x="Molarity", y="Line Scale Rating (cm)", color="Sweetener", shape="Sweetener") + scale_x_continuous(breaks=c(.1, .3, .5, .7, .9, 1.1, 1.3)) + theme_bw()
```

```{r, fig.width=10,fig.height=8}
# Q4: Molarity v. magnitude estimation
ggplot(line, aes(x=Mol, y=ME, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=MEMean-MESE, ymax=MEMean+MESE), size=1, width=.1, alpha=.2) + labs(title="Molarity Versus Magnitude Estimation", x="Molarity", y="Magnitude Estimaton", color="Sweetener", shape="Sweetener") + scale_x_continuous(breaks=c(.1, .3, .5, .7, .9, 1.1, 1.3)) + scale_y_continuous(breaks=c(1, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120))+ theme_bw()
```


```{r, fig.width=12,fig.height=8}
#Q5 Log mol v. line scale rating
ggplot(line, aes(x=LogMol, y=LS, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), size=1, width=.05, alpha=.2) + labs(title="Log Molarity Versus Line Scale Rating", x="Log Molarity", y="Line Scale Rating (cm)", color="Sweetener", shape="Sweetener") + theme_bw()
```

```{r, fig.width=12,fig.height=8}
#Q6 Log mol v. magnitude estimation
ggplot(line, aes(x=LogMol, y=ME, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=MEMean-MESE, ymax=MEMean+MESE), size=1, width=.05, alpha=.2) + labs(title="Log Molarity Versus Magnitude Estimation", x="Log Molarity", y="Magnitude Estimation", color="Sweetener", shape="Sweetener") + scale_y_continuous(breaks=c(1, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120)) + theme_bw()
```

```{r, fig.width=12,fig.height=8}
#Q7: Log molarity v. log line scale ratings
# Calculate log line scale ratings
line$logline <- log10(line$LS)
# Calculate new standard error
logse <- std.error(line$logline)

ggplot(line, aes(x=LogMol, y=logline, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=LogMean-logse, ymax=LogMean+logse), size=1, width=.05, alpha=.2) + labs(title="Log Molarity Versus Log Line Scale Rating", x="Log Molarity", y="Log Line Scale Rating (cm)", color="Sweetener", shape="Sweetener") + theme_bw()
```

```{r, fig.width=12,fig.height=8}
#Q8: Log molarity v. log magnitude estimation
# Calculate log line scale ratings
line$logme <- log10(line$ME)
# Calculate new standard error
logmese <- std.error(line$logme)

ggplot(line, aes(x=LogMol, y=logme, color=FS, shape=FS)) + geom_point(size=2) + geom_errorbar(aes(ymin=MELogMean-logmese, ymax=MELogMean+logmese), size=1, width=.05, alpha=.2) + labs(title="Log Molarity Versus Log Magnitude Estimation", x="Log Molarity", y="Log Magnitude Estimation", color="Sweetener", shape="Sweetener") + theme_bw()
```

Written Responses

1. When measured in weight-by-weight units, fructose appears sweeter- meaning, it takes less weight of fructose to achieve an equivalently sweet taste to fructose. However, sucrose is appears sweeter when measured by molarity, so it takes less molecules of sucrose to volume of product to achieve a sweetness equivalent to fructose. Looking at the graphs, especially the error bars relative to the y-axis means, I notice that fructose appears to have higher perceived sweetness at the highest weights/molarities- therefore, even though it may take greater molecular density of fructose to achieve a desired sweetness, that concentration will be perceived as sweeter than an equivalent percent concentration of sucrose.

2. When using weight-by-weight units, it is more clear that fructose is perceived as sweeter at equivalent percent concentrations. However, when considering molarity, it appears as though more fructose is needed to achieve similar sweetness to sucrose: when in fact, it is more molecular density that is needed rather than a greater quantity by weight of fructose. Using logged scales allows for the normalization of data such that extreme outliers do not contribute to skew in the data. Comparing the fructose molarity values to the lower sucrose values (knowing that these are equivalent in percent concentration) error bars shows a greater perceived sweetness. Logged value units make this easier to understand, though the unit different between molarity and weight-by-weight concentration could be of varying significance in different fields of study relating to sweeteners, whether it be in chemical analysis or purchasing evaluation of a sweetener.

3. The relative potency of fructose and sucrose is that fructose is sweeter per unit of the weight to volume ratio, but sucrose is sweeter per molecule to volume unit.

